---
title: "The Leslie Matrix - revisited"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

A population of fish we're interested in has four life stages: eggs (E), fry (F), juvenile (J), breeding adult (A). You are told that annually:

- Each breeding adult will produce, on average, 600 eggs
- 5% of eggs will survive to become fry
- 20% of fry will survive to become juveniles
- 16% of juveniles will survive to become adults
- Adult survival rate year to year is 90%

Recall the Leslie matrix from EDS 212. Matrix multiplication (using the dot product `%*%` of the matrix with the populations at each life stage) yields the population structure at the following year. 

### Create the Leslie matrix:

```{r}
# Make the Leslie Matrix: (byrow = TRUE means you're populating values across the rows, default is to populate down columns)
fish_leslie <- matrix(c(0, 0, 0, 600, 
                        0.05, 0, 0, 0, 
                        0, 0.20, 0, 0, 
                        0, 0, 0.16, 0.9), nrow = 4, ncol = 4, byrow = TRUE)

# Check it out: 
fish_leslie
```

The initial population structure is given by: 
```{r}
# Initial population structure (Roe, Fry, Juvenile, Adult): 
fish_year0 <- c(5000, 8000, 600, 400)
```


Add line-by-line comments wheres prompted in the code below, explaining what each is doing. Make sure that you are talking with your neighbors about what is really being created at each step. 

```{r}
# Model projections

# ------ WHAT HAPPENS HERE? ------ #
# ------ comment line-by-line ---- #

# The line below we are creating a vector of the time period that we want to run this for (8 years), telling it to go by increments of 1 year.
time_yr <- seq(from = 0, to = 8, by = 1)

# The line below creates an empty matrix with 4 columns to be populated by the output of the for loop below.
proj_year <- matrix(ncol = length(fish_year0), nrow = length(time_yr))

# The line below populates the first row (aka year 0) of the matrix (proj_year) with the initial conditions given above (fish_year0).
proj_year[1, ] <- fish_year0

# The for loop below starts at row 2 and does the dot product (aka multiplies matrices) based on the fish_leslie directions and the previous year's outcomes over the sequence period created (8 years).
for (i in 2:length(time_yr)) {
  proj_year[i,] <- fish_leslie %*% proj_year[i-1,]
}


# The rest is wrangling & visualization (run to check it out):
colnames(proj_year) <- c("eggs", "fry", "juvenile", "adult")

proj_df <- data.frame(time_yr, proj_year) %>% 
  pivot_longer(cols = -time_yr, names_to = "lifestage", values_to = "stage_population")

ggplot(data = proj_df, aes(x = time_yr, y = stage_population)) +
  geom_line(aes(color = lifestage)) +
  scale_y_log10()

```

Test different combinations of survival / reproduction and starting population structures to see what happens! 
